Technical Report

### 2018-04-25 fully-abstract-comp/technical-report.tex
added missing i in world specifies component.
Fix use of wrong macro in nsubeq def.
Small details.
Finish another supporting proof for the ftlr about stackreadcond and fix minor problems I noticed here and there while making the proof.

@@

\[
  \readCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \\
      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
    \end{array}
  \right\}
\]
\lau{RESOLVED: Make this condition require a set of adjacent regions. Check whether needed for the stack condition.}
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
\[
  \stackReadCond{W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \\
      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge \\
      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
    \end{array}
  \right\}
\]
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
\[
  \xReadCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
      \quad \exists M : S \fun \powerset{\MemSeg}, R : S \fun \powerset{\Seal}, C : S \fun \powerset{\Seal} \\
      \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr] \wedge\\
      % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
      \qquad \quad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
    \end{array}
  \right\}
\]
\lau{RESOLVED: Make this condition require a set of adjacent regions.}
\lau{TODO: allow code regions to be merged into one. Allow for (... forgot to write the rest, but probably related to splicing regions.) }

@@

\[
  \writeCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr)}) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \\
      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
      \qquad \quad \pwheap(r) \text{ is address-stratified}
    \end{array}
  \right\}
\]
\lau{TODO: Make this condition require a set of adjacent regions. Check whether necessary for the stack region.}
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
% \dominique{Why $n-1$ above?  There is already a quantification over $n' <n$ in the memory satisfaction relation and another $n-1$ is used in $\stdreg{[\baddr',\eaddr'],\gc}{\pur}$. Also: readCond doesn't have the $n-1$. Update: the memory satisfaction was updated by Lau to remove the quantification over $n' < n$ and I dropped the $n-1$ here.}
\[
  \stackWriteCond{W} = \left\{ \npair{(\baddr,\eaddr)}) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \\
      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge \\
      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsupeq[n-1] \stdreg{R(r),\gc}{\pur} \wedge\\
      \qquad \quad \pwfree(r) \text{ is address-stratified}
    \end{array}
  \right\}
\]
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.

@@

\subsection{Standard regions}
\label{sec:standard-regions}
Standard region:
\[
  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std},\square} \; \gc)
\]
for readability, we use $\spao$ short for $\spatialo$, $\spa$ short for $\spatial$, and $\pur$ as short for $\pure$.
\lau{02-11-2017: I am not sure whether it should be a $\normal$ region. In previous work $\nsubeq$ ignored the ``view'' which it may also do here, and then then it does not matter which one it is.}
where $H^{\mathrm{std},\square}_A$ is defined as follows:
\[
  H_A^{\mathrm{std},\square} \; \gc \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
    \begin{array}{l}
      \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
      \exists S : A \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
      \quad \forall \aaddr \in A, n' < n \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
    \end{array}
  \right\}
\]
\lau{01-11-2017: RESOLVED look into what kind of world separation should happen here.}
%\dominique{7-3-2018: is it necessary to have $n-1$ above? there is already a quantification over $n'<n$ in the memory satisfaction relation.} \lau{04-04-2018: The n'<n was removed from memory satisfaction.}

@@

\begin{equation*}
  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_) \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
    \begin{array}{l}
    \dom(\code) = [\aaddr,\baddr] \wedge \\
    \dom(\code) \nsubseteq \ta \Rightarrow \sigrets = \emptyset \wedge \\
    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code 
    \end{array}
  \right\}
\end{equation*}

@@

\begin{definition}[World specifies component]
  We define that a world $W$ specifies the presence of a component with global constants $\gc$
  \begin{equation*}
    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
  \end{equation*}
  (written as $W \vdash_\gc \var{comp}_0$)
  iff
  \begin{equation*}
    \vdash (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
  \end{equation*}
  and there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that
  \begin{itemize}
  \item $\pwheap(r_i) = \codereg{\sigrets_i,\sigcloss_i,\code_i,\gc}$
  \item $\mscode = \biguplus_i \code_i$
  \item One of the following holds:
    \begin{itemize}
    \item $\dom(\mscode) \subseteq \ta$ and $\sigrets = \biguplus_i \sigrets_i$ and $\sigcloss = \biguplus_i \sigcloss_i$.
    \item $\dom(\mscode) \mathrel{\#} \ta$ and $\sigrets = \emptyset$ and $\sigcloss = \biguplus_i \sigrets_i \uplus \biguplus_i \sigcloss_i$.
    \end{itemize}
  \end{itemize}
  and there exist $r_1 \cdots r_m \in \dom(\pwheap)$ such that
  \begin{itemize}
  \item $\pwheap(r_i) = \stdreg{A_i,\gc}{\pure}$
  \item $\dom(\msdata) = \biguplus_i A_i$
  \end{itemize}
  with $\gc = (\ta,\stkb)$
\end{definition}
Note that for untrusted components, we require all seals to be in the corresponding code region as closure seals.
That is because we cannot trust untrusted components to treat their seals correctly (e.g. not hand them out to other components, only sign return capabilities with return seals etc.), so instead, they must be considered as closure seals, which the component is essentially allowed to sign anything with.

>>>>

\dominique{Remaining question: what exactly does $\nsubeq$ on a UPred mean?  Does it mean subset ``up to n (including)'' or ``up to n-1 (including)''?
  Check what we can require without breaking contractiveness of recursive definition of value relation and/or what we did in previous paper and/or what we need in FTLR proof.
  Update: in the FTLR proof, we seem to need ``up to n'' because the heap predicate of the standard region is contractive in terms of the value relation.
  This also means we can afford to use ``up to n'' without breaking contractiveness of the recursive definition of the value relation.
  }
We have the following notion of regions being an $n$-subset of another region.
\begin{mathpar}
  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
            { (v,H) \nsubeq (v',H') }
\and
  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
            { (\pure,H,H_\sigma) \nsubeq (v',H') }
\and
  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
            { (v,H) \nsubeq (\pure,H',H_\sigma) }
\and
  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
            { (\pure,H,H_\sigma) \nsubeq (\pure,H',H_\sigma') }
\end{mathpar}
We define $n$-equality of regions $\iota \nequal \iota'$ as the symmetric closure, i.e. $\iota \nequal \iota'$ iff $\iota \nsubeq \iota'$ and $\iota \nsupeq \iota'$.

@@

\[
  \readCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \\
      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
    \end{array}
  \right\}
\]
\lau{RESOLVED: Make this condition require a set of adjacent regions. Check whether needed for the stack condition.}
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
\[
  \stackReadCond{W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \\
      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge \\
      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
    \end{array}
  \right\}
\]
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
\[
  \xReadCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
      \quad \exists M : S \fun \powerset{\MemSeg} \\
      \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr] \wedge\\
      % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
      \qquad \quad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
    \end{array}
  \right\}
\]
\lau{RESOLVED: Make this condition require a set of adjacent regions.}
\lau{TODO: allow code regions to be merged into one. Allow for (... forgot to write the rest, but probably related to splicing regions.) }

@@

\[
  \writeCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr)}) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \\
      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
      \qquad \quad \pwheap(r) \text{ is address-stratified}
    \end{array}
  \right\}
\]
\lau{TODO: Make this condition require a set of adjacent regions. Check whether necessary for the stack region.}
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
% \dominique{Why $n-1$ above?  There is already a quantification over $n' <n$ in the memory satisfaction relation and another $n-1$ is used in $\stdreg{[\baddr',\eaddr'],\gc}{\pur}$. Also: readCond doesn't have the $n-1$. Update: the memory satisfaction was updated by Lau to remove the quantification over $n' < n$ and I dropped the $n-1$ here.}
\[
  \stackWriteCond{W} = \left\{ \npair{(\baddr,\eaddr)}) \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \\
      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge \\
      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
      \qquad \quad \pwfree(r) \text{ is address-stratified}
    \end{array}
  \right\}
\]
where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.

@@

\subsection{Standard regions}
\label{sec:standard-regions}
Standard region:
\[
  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std},\square} \; \gc) , v \in \{\spa,\spao\}
\]
for readability, we use $\spao$ short for $\spatialo$, $\spa$ short for $\spatial$, and $\pur$ as short for $\pure$.
\[
  \stdreg{A,\gc}{\pur} \defeq (\pur,H_A^{\mathrm{std},\square} \; \gc, \lambda \_ \; \_ \ldotp \emptyset)
\]
\lau{02-11-2017: I am not sure whether it should be a $\normal$ region. In previous work $\nsubeq$ ignored the ``view'' which it may also do here, and then then it does not matter which one it is.}
where $H^{\mathrm{std},\square}_A$ is defined as follows:
\[
  H_A^{\mathrm{std},\square} \; \gc \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
    \begin{array}{l}
      \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
      \exists S : A \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
      \quad \forall \aaddr \in A, n' < n \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
    \end{array}
  \right\}
\]
\lau{01-11-2017: RESOLVED look into what kind of world separation should happen here.}
%\dominique{7-3-2018: is it necessary to have $n-1$ above? there is already a quantification over $n'<n$ in the memory satisfaction relation.} \lau{04-04-2018: The n'<n was removed from memory satisfaction.}

@@

\begin{equation*}
  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_) \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
    \begin{array}{l}
    \dom(\code) = [\aaddr,\baddr] \wedge \\
    \dom(\code) \subseteq \ta \Rightarrow \sigrets = \emptyset \wedge \\
    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code 
    \end{array}
  \right\}
\end{equation*}

@@

\begin{definition}[World specifies component]
  We define that a world $W$ specifies the presence of a component with global constants $\gc$
  \begin{equation*}
    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
  \end{equation*}
  (written as $W \vdash_\gc \var{comp}_0$)
  iff
  \begin{equation*}
    \vdash (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
  \end{equation*}
  and there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that
  \begin{itemize}
  \item $\pwheap(r_i) = \codereg[\mathrm{code}_i,\square]{\sigrets_i,\sigcloss_i,\code_i,\gc}$
  \item $\mscode = \biguplus_i \code_i$
  \item One of the following holds:
    \begin{itemize}
    \item $\dom(\mscode) \subseteq \ta$ and $\sigrets = \biguplus_i \sigrets_i$ and $\sigcloss = \biguplus_i \sigcloss_i$.
    \item $\dom(\mscode) \mathrel{\#} \ta$ and $\sigrets = \emptyset$ and $\sigcloss = \biguplus_i \sigrets_i \uplus \biguplus_i \sigcloss_i$.
    \end{itemize}
  \end{itemize}
  and there exist $r_1 \cdots r_m \in \dom(\pwheap)$ such that
  \begin{itemize}
  \item $\pwheap(r_i) = \stdreg{A_i,\gc}{\pure}$
  \item $\dom(\msdata) = \biguplus_i A_i$
  \end{itemize}
  with $\gc = (\ta,\stkb)$
\end{definition}
Note that for untrusted components, we require all seals to be in the corresponding code region as closure seals.
That is because we cannot trust untrusted components to treat their seals correctly (e.g. not hand them out to other components, only sign return capabilities with return seals etc.), so instead, they must be considered as closure seals, which the component is essentially allowed to sign anything with.
